Computationally Efficient High-Dimensional Regularized Discriminant Analysis

### 2017-01-23 reports/paper-hdrda.Rnw
Tightened up the discussion.

We have compared our proposed \emph{HDRDA} classifier to the random-forest and
Witten classifiers as well as the Pang, Tong, and Guo classifiers, each of which
assumes independence of the features conditional on class membership. Despite
the rapid computational performance of diagonal classifiers and the reduction in
the number of parameters to estimate, \cite{Mai:2012bf} and \cite{Fan:2012iq}
have noted that diagonal classifiers can often yield inferior classification
performance when compared to other classification methods. In fact, we have
shown that the \emph{HDRDA} classifier often yields superior classification
accuracy when compared to the diagonal classifiers considered using six
small-sample, high-dimensional data sets, confirming the assertions of
\cite{Mai:2012bf} and \cite{Fan:2012iq}. Furthermore, we have demonstrated that
\emph{HDRDA} classifiers are more robust to the presence of outliers than the
diagonal classifiers.

We also considered the popular penalized linear discriminant analysis from
\cite{Witten:2011kc} because it was specifically designed for high-dimensional
gene-expression data. We had expected its classification performance to be
competitive within our classification study and perhaps superior. Contrarily,
its underperformance in our empirical studies show that it is sensitive to
outliers and unable to achieve comparable results with other classifiers
designed for small-sample, high-dimensional data. Also, despite the claims of
\cite{FernandezDelgado:2014ul} that the random forest classifier is typically
superior to other classifiers, we observed that the random-forest classifier was
indeed competitive but was typically outperformed by classifiers developed for
small-sample, high-dimensional data.

We demonstrated that the \emph{HDRDA} implementation of
\eqref{eq:hdrda-decomposed} in the {\tt sparsediscrim} R package can be used in
practice with high-dimensional data sets. For instance, in our timing
comparisons we demonstrated that training and model selection could be employed
on data sets having 5,000 features in 3 seconds on average. We contrast our
implementation to the the poor computational performance of the in the {\tt
  klaR} R package, which took approximately 25 minutes on average to train the
\emph{RDA} classifier on a data set having 5,000 features. Given that the \emph{RDA}
classifier has been shown to have excellent performance in the high-dimensional
setting \citep{Webb:2011vu} but is limited by its computationally intense
model-selection procedure, our work enables the \emph{RDA} classifier to be
employed with small-sample, high-dimensional data in practice. This result is
reassuring because the \emph{RDA} classifier remains widely popular in the
literature. In fact, variants of the RDA classifier have been applied to
microarray data \citep{Ching:2012fu,Li:2012ev,Tai:2007bk,Guo:2007te}, facial
recognition
\citep{Ching:2012fu,Zhang:2010va,Dai:2007vd,Lu:2005hq,Pima:2004vw,Lu:2003we},
handwritten digit recognition \citep{Bouveyron:2007gx}, remote sensing
\citep{Tadjudin:1999fk}, seismic detection \citep{Anderson:2002kg}, and chemical
spectra \citep*{Wu:1996us,Aeberhard:1993fp}.

The dimension reduction employed in this paper has reduced the
dimension to rank$(\widehat{\bm \Sigma}) = q$. An interesting extension of our
work would reduce the dimension $q$ further to a lower dimension $q_L < q$, perhaps using a criterion similar to that of principal
components analysis. While unclear whether the classification performance would
improve via such a method, the efficiency of the training and model selection
would certainly improve. Moreover, if $q_L = 2$ or $3$, low-dimensional graphical displays of high-dimensional data could
be obtained.

>>>>

We have demonstrated that our proposed \emph{HDRDA} classifier is competitive
with and often superior to random forests as well as the Witten, Pang, Tong, and
Guo classifiers. In fact, we have shown that the \emph{HDRDA} classifier often
yields superior classification accuracy when applied to small-sample,
high-dimensional data sets, confirming the assertions of \cite{Mai:2012bf} and
\cite{Fan:2012iq} that diagonal classifiers often yield inferior classification
performance when compared to other classification methods. Furthermore, we have
demonstrated that \emph{HDRDA} classifiers are more robust to the presence of
outliers than the diagonal classifiers despite their rapid computational
performance and their reduction in the number of parameters to estimate.

We also considered the popular penalized linear discriminant analysis from
\cite{Witten:2011kc} because it was specifically designed for high-dimensional
gene-expression data. We had expected its classification performance to be
competitive within our classification study and perhaps superior. Contrarily,
our empirical studies suggests that the classifier is sensitive to outliers and
unable to achieve comparable results with other classifiers designed for
small-sample, high-dimensional data. Also, despite the claims of
\cite{FernandezDelgado:2014ul} that random forests are typically superior to
other classifiers, we observed that they were indeed competitive but were
typically outperformed by classifiers developed for small-sample,
high-dimensional data.

We demonstrated that our \emph{HDRDA} implementation in the {\tt sparsediscrim}
R package can be used in practice with high-dimensional data sets. In our timing
comparisons we showed that \emph{HDRDA} model selection could be employed on
data sets with $p = 5000$ in \Sexpr{hdrda_mean_time_p5000} seconds on
average. Contrarily, the \emph{RDA} classifier implemented in the {\tt klaR} R
package required \Sexpr{rda_mean_time_p5000} minutes on average to perform model
selection on data sets with $p = 5000$. Given that the \emph{RDA} classifier has
been shown to have excellent performance in the high-dimensional setting
\citep{Webb:2011vu} but is limited by its computationally intense
model-selection procedure, our work replaces the \emph{RDA} classifier for
high-dimensional data in practice. This result is reassuring because the
\emph{RDA} classifier remains widely popular in the literature. In fact,
variants of the RDA classifier have been applied to microarray data
\citep{Ching:2012fu,Li:2012ev,Tai:2007bk,Guo:2007te}, facial recognition
\citep{Zhang:2010va,Dai:2007vd,Lu:2005hq,Pima:2004vw,Lu:2003we},
handwritten digit recognition \citep{Bouveyron:2007gx}, remote sensing
\citep{Tadjudin:1999fk}, seismic detection \citep{Anderson:2002kg}, and chemical
spectra \citep{Wu:1996us,Aeberhard:1993fp}.

The dimension reduction employed in this paper has reduced the dimension to
rank$(\widehat{\bm \Sigma}) = q$. An interesting extension of our work would
reduce the dimension $q$ further to a lower dimension $q_L < q$, perhaps using a
criterion similar to that of principal components analysis. While unclear
whether the classification performance would improve via such a method, the
efficiency of the model selection would certainly improve. Moreover, if $q_L =
2$ or $3$, low-dimensional graphical displays of high-dimensional data could be
obtained.
